---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document: 
    keep_md: true
---


## Loading and preprocessing the data
````{r}
df <- read.csv(unz("activity.zip", "activity.csv"), header=T, sep="," )
df$date <- as.Date(df$date)
dim(df)
summary(df)


```


## What is mean total number of steps taken per day?

````{r}
library(dplyr)
dailySteps <- df %>%
  group_by(date)%>%
  summarise(sumSteps = sum(steps, na.rm = TRUE))
#dailySteps
hist(dailySteps$sumSteps, breaks = 25,
     main = "Total number of steps per day",
     xlab = "Total daily steps",
     col = "green",
     xlim = c(0,25000))
````


### Calculate and report the mean and median of the total number of steps taken per day
````{r}
totalStepsMean <- mean(dailySteps$sumSteps, na.rm = T)
totalStepsMedian <- median(dailySteps$sumSteps, na.rm = T)

print(paste("The mean of the total number of steps taken per day is",totalStepsMean ,"and the median is" ,totalStepsMedian ,"."))

````

## What is the average daily activity pattern?

````{r}
dailyActivity <- as.data.frame(df %>%
    group_by(interval) %>%
    summarise(avgActivity = mean(steps, na.rm = TRUE)))

plot(dailyActivity$interval, dailyActivity$avgActivity, 
     type = "l", 
     xlab = "Interval", 
     ylab= "Mean of Steps Across All Days", 
     main = "Daily Activity Pattern")


````


### Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
````{r}
dailyActivity[dailyActivity[,"avgActivity"] == max(dailyActivity$avgActivity),]$interval
````

## Imputing missing values


###Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)

````{r}
table(is.na(df))["TRUE"]

````


### Devise a strategy for filling in all of the missing values in the dataset. The strategy does not need to be sophisticated
````{r}
library(mice)
tempData <- mice(df,m=5,maxit=50,meth='pmm',seed=500)
summary(tempData)

````


### Create a new dataset that is equal to the original dataset but with themissing data filled in.

````{r}
completedDf <- complete(tempData,1)

````


### Make a histogram of the total number of steps taken each day and Calculate
### and report the mean and median total number of steps taken per day. Do
### these values differ from the estimates from the first part of the assignment?
### What is the impact of imputing missing data on the estimates of the total
### daily number of steps?###
````{r}
dailySteps2 <- completedDf %>%
  group_by(date)%>%
  summarise(sumSteps = sum(steps, na.rm = TRUE))

hist(dailySteps2$sumSteps, breaks = 25,
     main = "Total number of steps per day",
     xlab = "Total daily steps",
     col = "green",
     xlim = c(0,25000))

totalStepsMean2 <- mean(dailySteps2$sumSteps, na.rm = T)
totalStepsMedian2 <- median(dailySteps2$sumSteps, na.rm = T)

print(paste("The mean of the total number of steps taken per day is",totalStepsMean2 ,"and the median is" ,totalStepsMedian2 ,"."))
````

## Are there differences in activity patterns between weekdays and weekends?

````{r}
library(ggplot2)
library(lattice)
week <- c("Mon","Tue","Wed","Thu","Fri")

df1 <- completedDf %>%
  mutate(wday = weekdays(date, abbreviate = TRUE), wday = factor((wday %in% week), levels = c(FALSE, TRUE), labels=c('weekend', 'weekday'))) %>%
  group_by(date, interval, wday) %>%
  summarise(avg = mean(steps), date, wday, .groups = "keep")
  
xyplot(avg~interval|wday, df1, type='l', lty = 1, lwd = 0.5, ylab = "Number of steps", layout = c(1,2))
 

  


````
